### 2021-01-26 priority-gv.tex
Made a bunch of minor changes?
In the process of adding examples.

@@

The only non-standard feature of the typing rules is the priority annotations. Priorities are based on \emph{obligations/capabilities} used by Kobayashi~\cite{kobayashi06}, and simplified to single priorities following Padovani~\cite{padovani14}. The integration of priorities into GV is adapted from Padovani and Novara~\cite{padovaninovara15}. Paraphrasing Dardha and Gay~\cite{dardhagay18}, priorities obey the following two laws:
\begin{enumerate*}[label=(\roman*)]
\item an action with priority $\cs{o}$ happens before an action with priority $\cs{o+n}$; and
\item communication requires \emph{equal} priorities for the dual actions.
\end{enumerate*}

In PGV, sequents are annotated with a priority bound $\cs{p}$, which is the \emph{upper bound} on the priority of actions performed while evaluating the term. The \emph{lower bound} is obtained from the typing environment. Terms with sequential evaluation, such as \LabTirName{T-Andthen}, have side conditions to ensure deadlock freedom. For instance, the side condition for typing $\tm{\andthen{M}{N}}$ requires that $\tm{M}$ has finished communicating before $\tm{N}$ starts communicating. The rule \LabTirName{T-Lam} stores the lower and upper priority bound of the function body in the type.

>>>>

\todo{Describe examples.}

\begin{center}
\begin{minipage}{0.5\linewidth}
\begin{example}[Cyclic Structure]
  \[\tm{%
    \begin{array}{l}
      \letpair{x}{x'}{\new}{}
      \\
      \letpair{y}{y'}{\new}{}
      \\
      \spawn\left(%
      \begin{array}{l}
        {\andthen{\recv\;{x'}}{\send\;{\pair{\unit}{y}}}};
        \\
        {\andthen{\wait\;{x'}}{\close\;{y}}}
      \end{array}%
      \right);
      \\
      \underline{\andthen{\send\;{\pair{\unit}{x}}}{\recv\;{y'}}};
      \\
      {\andthen{\close\;{x}}{\wait\;{y'}}}
    \end{array}
  }\]
\end{example}
\end{minipage}%
\begin{minipage}{0.5\linewidth}
\begin{example}[Deadlock]
  \[\tm{%
    \begin{array}{l}
      \letpair{x}{x'}{\new}{}
      \\
      \letpair{y}{y'}{\new}{}
      \\
      \spawn\left(%
      \begin{array}{l}
        {\andthen{\recv\;{x'}}{\send\;{\pair{\unit}{y}}}};
        \\
        {\andthen{\wait\;{x'}}{\close\;{y}}}
      \end{array}%
      \right);
      \\
      \underline{\andthen{\recv\;{y'}}{\send\;{\pair{\unit}{x}}}}
      \\
      {\andthen{\close\;{x}}{\wait\;{y'}}}
    \end{array}
  }\]
\end{example}
\end{minipage}
\end{center}

@@

The only non-standard feature of the typing rules is the priority annotations. Priorities are based on \emph{obligations/capabilities} used by Kobayashi~\cite{kobayashi06}, and simplified to single priorities following Padovani~\cite{padovani14}. The integration of priorities into GV is adapted from Padovani and Novara~\cite{padovaninovara15}. Paraphrasing Dardha and Gay~\cite{dardhagay18}, priorities obey the following two laws:
\begin{enumerate*}[label=(\roman*)]
\item an action with lower priority happens before an action with higher priority; and
\item communication requires \emph{equal} priorities for dual actions.
\end{enumerate*}

In PGV, we keep track of a lower and upper bound on the priorities of a term, \ie, while evaluating the term, when does it start communicating, and when does it finish. The upper bound is written on the sequent, whereas the lower bound is approximated from the typing environment. Typing rules for sequential constructs enforce sequentially, \eg, the typing for $\tm{\andthen{M}{N}}$ has a side condition which requires that the upper bound of $\tm{M}$ is smaller than the lower bound of $\tm{N}$, \ie, $\tm{M}$ finishes before $\tm{N}$ starts.
\begin{mathpar}
  \inferrule*{
    \inferrule*{
      \inferrule*{
      }{
        \tseq{\tmty{y'}{\tyrecv{\tyunit}{\tyendr}}}{y'}{\tyrecv{\tyunit}{\tyendr}}
      }
    }{
      \tseq{\tmty{y'}{\tyrecv{\tyunit}{\tyendr}}}{\recv\;{y'}}{\typrod{\tyunit}{\tyendr}}
    }
    \\
    \inferrule*{
    }{
      \tseq{}{\send\;{\pair{\unit}{x}}}{}
    }
  }{
    \tseq{}{\andthen{\send\;{\pair{\unit}{x}}}{\recv\;{y'}}}{}
  }

  {\andthen{\recv\;{y'}}{\send\;{\pair{\unit}{x}}}}
\end{mathpar}
Closures suspend communication, so \LabTirName{T-Lam} stores the priority bounds of the function body on the function type, and \LabTirName{T-App} restores them. For instance, $\tm{\lambda{x}.\send\;\pair{x}{y}}$ is assigned the type $\ty{\tylolli[\cs{o},\cs{o}]{A}{S}}$, \ie, a~function which, when applied, starts and finishes communicating at priority $\cs{o}$:
\begin{mathpar}
  \inferrule*{
    \inferrule*{
      \inferrule*{}{\tm{\send}}
      \\
      \inferrule*{
        \inferrule*{
        }{
          \tseq[\pbot]{\tmty{x}{A}}{x}{A}
        }
        \\
        \inferrule*{
        }{
          \tseq[\pbot]
          {\tmty{x}{A},\tmty{y}{\tysend[\cs{o}]{A}{S}}}
          {y}{\tysend[\cs{o}]{A}{S}}
        }
      }{
        \tseq[\pbot]
        {\tmty{x}{A},\tmty{y}{\tysend[\cs{o}]{A}{S}}}
        {\pair{x}{y}}{\typrod{A}{\tysend[\cs{o}]{A}{S}}}
      }
    }{
      \tseq[\cs{o}]
      {\tmty{x}{A},\tmty{y}{\tysend[\cs{o}]{A}{S}}}
      {\send\;\pair{x}{y}}{S}
    }
  }{
    \tseq[\pbot]
    {\tmty{y}{\tysend[\cs{o}]{A}{S}}}
    {\lambda{x}.\send\;\pair{x}{y}}{\tylolli[\cs{o},\cs{o}]{A}{S}}
  }
\end{mathpar}
The approximation of the lower bound is not entirely accurate, \eg, $\tm{\lambda{x}.\send\;\pair{x}{y}}$ has lower bound $\ptop$, not $\cs{o}$, but as we only ever under-approximate, it is safe.
